How to be a good member of a scientific software community [Article v\versionnumber]

### 2022-01-05 response-letter.tex

\documentclass[11pt]{article}

\usepackage[numbers,super,sort&compress,comma]{natbib}
\usepackage[pdftex,colorlinks=true,urlcolor=black,filecolor=black,linkcolor=black,citecolor=black,breaklinks=true]{hyperref}
\usepackage{hypernat}
\usepackage[pdftex]{graphicx}
\usepackage{palatino} % make fonts Helvetica for sans, palatino for serif
\usepackage[usenames,dvipsnames]{color}

\newcommand{\response}[1]{\textcolor{Mahogany}{\textbf{#1}}}

\begin{document}

I thank the reviewers for their kind and constructive comments. I hope they agree that incorporating their suggestions has significantly improved the manuscript. I've listed the comments requiring responses below, with my response colored \response{in mahogony}.

\section{Reviewer A}

1) My biggest concern is simply that the people who ought to read it,
won't do so (even if directed) because it is longer than what users
typical want to read when they are in search of a fix to their
problem. In this the article shares the same shortcoming with the
well-known online essays [1] and [2].

I'd like a condensed summary of about one paragraph (maybe as a
box) that I can just paste in an email or put on a website,
together with a link to the full version.

\response{I have added references to the Raymond and Tatham essays -- I'm
somewhat embarrassed I didn't think to add them myself, since they clearly
influenced my thinking when I read them. I would argue that the checklists make
a good short punchy summary, as the reviewer suggested. However, I agree having
a plain-text version wouldn't hurt, so I've added a ``Summary'' paragraph before
the checklists.}

2) Does the author make a distinction between software that is
advertised as useable (e.g. through project websites, presence of
forums, issue tracker, ...) and software that has been made
available (archive, repository, supplementary material)?

Can and should a user assume that a developer of the latter will
provide any support?

Perhaps this could be made clearer.

My personal view is that FAIR principles of research make it
essential that code is published together with the research (code
dump with a license). However, for code dumps I wouldn't require
authors to respond to more than the most fundamental inquiries
(e.g., to respond when bugs are found that question the integrity
of the published research). If more is demanded of authors then
that disincentivizes publishing the code with the research.

On the other hand, if code is advertised (presumably with the
expectation of citations or some other benefit to the developer)
then I agree with the author that developers owe their users some
of their time.

\response{My focus was mostly on code that is intended to be used. I've added a
paragraph at the end of the section for developers laying out my thoughts on
this distinction.}

3) "Please open a PR" (or merge request) is a phrase that can strike
fear in the heart of a novice. It might be helpful to demystify
this important part of collaborative software development. Maybe
the a PLoS Comp Biol article [5] or something similar that touches
upon basic software engineering that could provide more guidance.

\response{I agree, and I'd argue that the relevance here is that it falls to the
developers to provide a less intimidating way to contribute. I'd rather not talk
about specific software engineering technologies here, though, just because I
feel like it's a side topic I couldn't do justice to.}

4) It might be useful to prepare the new community member that
conversations — although respectfule — are often to the point and
that most developers won't sugar-coat their words: if something is
wrong, it will be said just like that because ultimately developers
are concerned with maintaining the health of the code base and the
project and have to be clear about what is acceptable.

New users should not feel offended by the phrase "Feel free to work
on it, submit a PR, and we can discuss there." or similar. Unless
you pay the developers to implement what you want, they will likely
encourage you to do most of the work yourself and then spend some
of their scarce time guiding you to a solution.

\response{This is a good point (although I'd argue that developers ought to be a bit more courteous than that). I added a couple of sentences at the end of the bug report section, appended to the paragraph about feature requests.}

5) Pointing out how much one can learn by interacting with experienced
developers is important. This is a chance to learn from the best
people in the field.

Furthermore, the networks that you build as part of a software
project might last you a life time.

\response{I added a sentence mentioning this to the section on contributing to the community.}

6) I appreciate that the article primarily originates in the
experience of the author. I would nevertheless find it useful if,
where possible, other sources could also be cited as validation and
to show that indeed these are broadly shared values.

I found the following references useful but I am emphatically not
implying that they must be included in the article.

Ref [3] has a good section on the importance of communities.

\response{I'm embarrassed to admit I hadn't read either of the suggested papers.
I've now read them and added references to both.}

\section{Reviewer B}

1) This point is already implicitly transparent throughout the paper, but it
might be useful to state this more explicitly right in Section 3.1 as it
(partially) motivates everything that follows: The easier you make it for the
developer to narrow down/reproduce the issue the higher the chance they will
help you. I think it's worth pointing out that developers are usually handling
multiple issues at the same time and typically tend to assign lower priority to
issues that are harder and time-consuming.

\response{This is a very good point. I added a new paragraph under the description of a bug report.}

2) In section 3.5 it is stated: "Once you’ve become reasonably competent with a
piece of software, it’s worth looking for ways to contribute back. [...] You can
work on the documentation". This is a wonderful suggestion. However, I'm afraid
that stated this way, users that don't feel competent enough will be
hesitant/justified in contributing. I believe proposing small changes to the
documentation often does not require competency. On the contrary, devs and
experienced users typically take for granted a lot of things that might be
missing from the docs. My point is that it might be a good idea to rephrase a
bit to encourage even first users to propose changes to the docs as the best way
to give back some of the time the devs/community has dedicated to them.

\response{This is a very good point. The new version begins ``Once you've begun using a piece of software, it's worth looking for ways to
contribute back.  This can take many forms, not all of which involve writing
code (though of course contributing bug fixes or new features is extremely
valuable). You don't have to be particularly expert -- sometimes the best
documentation comes from the inexperienced or newly experienced users, who
vividly remember not understanding something in the manual and can suggest
revisions that would make it clearer. ''}

\section{Reviewer C}

\response{Thank you for your kind words!}

\end{document}

### 2022-01-05 paper.tex
updated COI
major changes suggested by reviewers and wrote response letter

The next step in the process is to actually communicate your issue with the
developers, generally in the form of a bug report.  Doing so effectively
requires striving for two somewhat contradictory goals: you must give the
developer all of the relevant information about the problem, as concisely as
possible.

@@

@@

Feature requests are similar: you need to describe what you hope to accomplish
and how you'd like to see it work, as well as any current workaround you've
developed. Keep in mind that many features sound easy, but there can be deep
design issues that make their implementation difficult. Moreover, it is possible
that the developer won't be enthusiastic about your idea, even if it is doable
--- for example, it might not coincide with their vision for the software. Or,
as discussed above, academic developers are generally not compensated for their
work (obtaining NIH funds to support software development is extremely
difficult) and simply might not have the time to work on new features.

@@

Once you've become reasonably competent with a piece of software, it's worth
looking for ways to contribute back. This can take many forms, not all of which
involve writing code (though of course contributing bug fixes or new features is
extremely valuable). You can work on the documentation --- if you struggled to
understand something but eventually figured it out, consider contributing a new
paragraph of text. Many packages hosted on GitHub have the documentation in the
same repository, so the same pull request mechanism can be used, but even if the
docs are separate you could write a new paragraph or two and send them to the
developer. If you developed a toy example as you were figuring out how to use
the software, consider sharing it. If one of the provided examples had to be
tweaked to work, share the tweaks. If the software has an online forum (e.g.
mailing list, GitHub discussions or issues, etc), you might be able to
contribute by answering questions from less experienced users.

Unlike the other recommentations, this one isn't mandatory --- it's a ``nice to
have'' rather than a ``must have''. However, there are a number of advantages to
doing it. First and foremost, you'll be making a package that's valuable to you
stronger, which will help you with your own scientific projects. However,
participating productively in a software community is also a good way to build
your scientific communication skills. Finally, helping others is an excellent
way for a relatively junior scientist to build their reputation in the
community, which in turn can lead to future collaborations or even job
opportunities. At the very least, it's always valuable when a lot of people in
your field have positive associations with your name.

@@

Doing these things is valuable from a practical perspective; making it easy for
people to interact with us efficiently, even if it costs us some effort up
front, will make those interactions more pleasant. Even setting that aside, it's
the right way to do science --- FAIR principles for software development
\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
That means documenting it well and supporting people who try to use it.

@@

@@

@@

Alan Grossfield serves as a consultant to two companies: Moderna, Inc.
and Atelerix Life Sciences.

>>>>

The next step in the process is to actually communicate your issue with the
developers, generally in the form of a bug report.  The challenges in writing an
effective bug report have been discussed before in the context of free software
\cite{Raymond, Tatham}.  Doing so requires striving for two somewhat
contradictory goals: you must give the developer all of the relevant information
about the problem, as concisely as possible.

@@

Following this format (or something similar) makes it much easier for the
developer to diagnose your problem, and the easier you make it the more likely
they are to solve your problem quickly. While it's not ideal behavior, busy
developers (particularly those for whom software support is a side activity) may
shy away from starting work on a problem that looks hard or uninviting, while a
clear concise description of the problem is more likely to spur immediate
action.

@@

Feature requests are similar: you need to describe what you hope to accomplish
and how you'd like to see it work, as well as any current workaround you've
developed. Keep in mind that many features sound easy, but there can be deep
design issues or technical nuances that make their implementation difficult.
Moreover, it is possible that the developer won't be enthusiastic about your
idea, even if it is doable --- for example, it might not coincide with their
vision for the software. Or, as discussed above, academic developers are
generally not compensated for their work (obtaining NIH funds to support
software development is extremely difficult) and simply might not have the time
to work on new features. For these reasons, do not be surprised or insulted if
the developers suggest you work on a particular solution yourself and submit a
pull request. This doesn't mean they dislike your idea, just that they're busy -- they'll help guide you as best they can, but they can't do the work themselves.

@@

Once you've begun using a piece of software, it's worth looking for ways to
contribute back.  This can take many forms, not all of which involve writing
code (though of course contributing bug fixes or new features is extremely
valuable). You don't have to be particularly expert -- sometimes the best
documentation comes from the inexperienced or newly experienced users, who
vividly remember not understanding something in the manual and can suggest
revisions that would make it clearer.  Many packages hosted on GitHub have the
documentation in the same repository, so the same pull request mechanism can be
used, but even if the docs are separate (or if you find the idea of a pull
request intimidating) you could write a new paragraph or two and send them to
the developer. If you developed a toy example as you were figuring out how to
use the software, consider sharing it. If one of the provided examples had to be
tweaked to work, share the tweaks. If the software has an online forum (e.g.
mailing list, GitHub discussions or issues, etc), you might be able to
contribute by answering questions from less experienced users.

Unlike the other recommendations, this one isn't mandatory --- it's a ``nice to
have'' rather than a ``must have''. However, there are a number of advantages to
doing it. First and foremost, you'll be making a package that's valuable to you
stronger, which will help you with your own scientific projects, while getting
feedback from leading experts in the field. However, participating productively
in a software community is also a good way to build your scientific
communication skills. Finally, helping others is an excellent way for a
relatively junior scientist to build their reputation in the community, which in
turn can lead to future collaborations or even job opportunities. At the very
least, it's always valuable when a lot of people in your field have positive
associations with your name.

@@

Doing these things is valuable from a practical perspective, as discussed
elsewhere \cite{Heister-2013, Laayouni-2016}; making it easy for people to
interact with us efficiently, even if it costs us some effort up front, will
make those interactions more pleasant. Even setting that aside, it's the right
way to do science --- FAIR principles for software development
\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
That means documenting it well and supporting people who try to use it.

@@

Sometimes, code is published that is not particularly intended for use. For
example, it is good practice to release the scripts used to analyze data in
connection with the paper publishing the data, but many of these scripts are
one-offs, and the authors have little incentive to polish them to the point of
general useability. Rather, the scripts are essentially there as documentation,
not turn-key programs for others to use. In this case, the authors have less of
an obligation to provide support, although at a minimum laying out the software
necessary to run it (e.g. versions of python and numpy, or a Makefile and
compiler version) should be required.  That said, we recognize that some bit rot
is expected with academic code; once the student who wrote the code graduates
and moves on, it's very possible that there is no one else to maintain it.
Solving the long-term support and maintenance problem for academic software is a
major problem, well beyond the scope of this manuscript.

\section{Summary}

Given the importance of software to modern science and the challenges in writing
high quality code, virtually all working scientists will need to ask for help
from time to time. The keys to doing so effectively are remarkably simple: ask
your questions succinctly and only after putting in some effort to solve the
problem yourself, be considerate of other people's time and effort, and look for
opportunities to contribute back to the community. Developers should make it
easy for users to behave this way by providing clear documentation (including
the appropriate venues to ask for help), encouraging participation, and
insisting on respectful behavior from users and developers alike.

@@

The cover image was created by artistfymo (\url{https://linktr.ee/ArtistFrmYO}).

@@

Alan Grossfield serves as a consultant to two companies: Moderna, Inc.
and Atelerix Life Sciences, and is a shareholder in Atelerix Life Sciences.
